Evaluating infectious disease forecasts with allocation scoring rules

### 2023-07-27 alloscore_manuscript/allocation.tex
add some stuff related to exponential distributions example

\subsection{Allocation Scoring}

In a nutshell, the allocation score of a multivariate forecast $F$ is the unnecessary realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to $F$,
\begin{align}
\sum_i E_{F_i}[\text{unmet need in location $i$}]
\end{align}
is minimized, where the $F_i$ are the marginal distributions of $F$. By ``unnecesary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation 
in various locations are as close to each other as possible (see methods).  This will lead to the allocations provided by $F$ being quantiles of $F_i$ for some \emph{single} probability level $\tau$.

@@

The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S. Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals. WIS is a discrete approximation to CRPS, and can be obtained by using a distribution $p$ that has point masses at the probability levels corresponding to the endpoints of a finite set of prediction intervals.

@@

We now develop a scoring rule for probabilistic forecasts that measures the value of a forecast as an input to decision making about how to allocate limited resources to meet demand across multiple locations. As a concrete example, we take the resource to be a good such as ventilators or oxygen supply. An administrator is tasked with determining where to send these resources so as to meet demand among hospital patients in different facilities or states. In contrast to the decision-making problem in the previous section, the administrator is not able to control the total amount of supply; rather, their task is to determine how to allocate the fixed supply to different locations.

In this decision-making setting, an action $\mathbf{x} = (x_1, \ldots, x_n)$ is a vector specifying the amount that is allocated to each of the $n$ locations. We require that each $x_i$ is non-negative and that the total allocation across all locations does not exceed a constraint $K$ on the total available resources: $\sum_{i=1}^n x_i \leq K$. \elr{Since we got rid of the C parameter, maybe we should just make this a hard constraint here, $\sum_i x_i = K$?} The set $\mathcal{X}$ collects all possible allocations that satisfy these constraints. The eventually realized resource demand in each location is denoted by $\mathbf{y} = (y_1, \ldots, y_n)$. Again, these levels of demand are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_n)$ where $Y_i$ represents the as-yet-unknown level of resource demand in location $i$. Forecasts of demand in each location are collected in $F = (F_1, \ldots, F_n)$. We assume that the forecasts do not allow for the possibility of negative demand, i.e. the support of each $F_i$ is a subset of $\mathbb{R}^+$. As in the previous section, we assume that a loss $L$ is incurred for each unit of unmet need.

>>>>

\subsection{Overview of Allocation Scoring}

The allocation score of a multivariate forecast $F$ is the unnecessary realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to $F$,
\begin{align}
\sum_{i=1}^n E_{F_i}[\text{unmet need in location $i$}]
\end{align}
is minimized, where the $F_i$ are the marginal distributions of $F$. By ``unnecessary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation 
in various locations are as close to each other as possible (see methods).  This will lead to the allocations provided by $F$ being quantiles of $F_i$ for some \emph{single} probability level $\tau$.

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$ and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, with $\sigma_1 = 1 and $\sigma_2 = 5$. If we have $K = 5$ units of our resource available, the optimal allocation according to $F$ would be 4 units in location 1 and 6 units in location 2. This allocates to both locations at their forecast medians.  If, on the other hand, we only have $K = 3$ units available, we will necessarily have to allocate in location 2 at a quantile for $\tau = 0$, and therefore all 3 would go to location 2 since allocating any resources to location 1 would lead to unbalanced probabilities of need exceeding allocation.

suppose we have $n = 2$ locations, with a fixed total of $K=5$ units of available resources. to the forecasts $Y_1 \sim $ with $\sigma_1 = 1$ and $Y_2 \sim Exp(1 / \sigma_2)$ with $\sigma_2 = 5$.

@@

The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S. COVID-19 Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals. WIS is a discrete approximation to CRPS, and can be obtained by using a distribution $p$ that has point masses at the probability levels corresponding to the endpoints of a finite set of prediction intervals.

@@

We now develop a scoring rule for probabilistic forecasts that measures the value of a forecast as an input to decision making about how to allocate limited resources to meet demand across multiple locations. As a concrete example, we take the resource to be a good such as ventilators or oxygen supply. An administrator is tasked with determining where to send these resources so as to meet demand among hospital patients in different facilities or states. In contrast to the decision-making problem in the previous section, the administrator is not able to control the total amount of supply; rather, their task is to determine how to allocate the fixed supply to different locations. Following the structure of the discussion for the quantile score and CRPS, we first discuss the set up for a single value of the resource constraint $K$, which plays a role that is analogous to the cost/loss ratio for the quantile score, and then we discuss the possibility of integrating (or averaging) across values of $K$.

\subsubsection{The allocation score for a fixed resource constraint}

In the decision-making setting that we consider, an action $\mathbf{x} = (x_1, \ldots, x_n)$ is a vector specifying the amount that is allocated to each of $n$ locations. We require that each $x_i$ is non-negative and that the total allocation across all locations does not exceed a constraint $K$ on the total available resources: $\sum_{i=1}^n x_i \leq K$. \elr{Since we got rid of the C parameter, maybe we should just make this a hard constraint here, $\sum_i x_i = K$?} The set $\mathcal{X}$ collects all possible allocations that satisfy these constraints. The eventually realized resource demand in each location is denoted by $\mathbf{y} = (y_1, \ldots, y_n)$. Again, these levels of demand are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_n)$ where $Y_i$ represents the as-yet-unknown level of resource demand in location $i$. Forecasts of demand in each location are collected in $F = (F_1, \ldots, F_n)$. We assume that the forecasts do not allow for the possibility of negative demand, i.e. the support of each $F_i$ is a subset of $\mathbb{R}^+$. As in the previous section, we assume that a loss $L$ is incurred for each unit of unmet need.
